---
title: "Final Project"
author: "Teah Thies"
date: "2024-05-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data

```{r}
library(readr)
df <- read_csv("~/Desktop/clean_df.csv", show_col_types = FALSE)
ncol(df)
```

## Data Visualization - Wound Score

```{r}
wound_data <- df[, c("WoundScore_Appearance", "WoundScore_Size", "WoundScore_Perfusion", "WoundScore_Total")]
```

```{r}
library(ggplot2)
library(magrittr)

## WoundScore_Appearance
aggregate(WoundScore_Total ~ WoundScore_Appearance, wound_data, mean) %>%
  ggplot(aes(x = WoundScore_Appearance, y = WoundScore_Total)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Average WoundScore_Total by WoundScore_Appearance",
       x = "WoundScore_Appearance",
       y = "Average WoundScore_Total")
```

```{r}
## WoundScore_Size
aggregate(WoundScore_Total ~ WoundScore_Size, wound_data, mean) %>%
  ggplot(aes(x = WoundScore_Size, y = WoundScore_Total)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(title = "Average WoundScore_Total by WoundScore_Size",
       x = "WoundScore_Size",
       y = "Average WoundScore_Total")
```

```{r}
## WoundScore_Perfusion
aggregate(WoundScore_Total ~ WoundScore_Perfusion, wound_data, mean) %>%
  ggplot(aes(x = WoundScore_Perfusion, y = WoundScore_Total)) +
  geom_bar(stat = "identity", fill = "green") +
  labs(title = "Average WoundScore_Total by WoundScore_Perfusion",
       x = "WoundScore_Perfusion",
       y = "Average WoundScore_Total")
```

## Data Visualization - Patient

```{r}
patient_data <- df[, c("Age", "Ethnicity", "Gender")]
```

```{r}
# Function to create pie chart for each group
create_pie_chart <- function(data, variable, title) {
  color_palette <- rainbow(length(unique(data[[variable]])))
  variable_counts <- table(data[[variable]])
  pie(variable_counts, labels = paste(names(variable_counts), "-", variable_counts), 
      col = color_palette,
      main = title)
}

# Create pie charts for each variable
par(mfrow=c(1, 2))  # Arrange plots in one row with three columns
create_pie_chart(patient_data, "Ethnicity", "Ethnicity Distribution")
create_pie_chart(patient_data, "Gender", "Gender Distribution")
```

```{r}
## histogram for Age
hist(patient_data$Age, 
     col = "skyblue",   
     main = "Age Distribution",  
     xlab = "Age",          
     ylab = "Frequency",  
     border = "black",     
     breaks = 10)          
```

```{r}
hist(df$GoalScore_Compliance, 
     col = "purple",   
     main = "GoalScore_Compliance",  
     xlab = "Score",          
     ylab = "Frequency",  
     border = "black",     
     breaks = 10) 

hist(df$GoalScore_Total, 
     col = "orange",   
     main = "GoalScore_Motivation",  
     xlab = "Score",          
     ylab = "Frequency",  
     border = "black",     
     breaks = 10)

hist(df$GoalScore_Comprehension, 
     col = "purple",   
     main = "GoalScore_Comprehension",  
     xlab = "Score",          
     ylab = "Frequency",  
     border = "black",     
     breaks = 10)
```

## Cohen's Kappa

```{r}
cohen_data <- df[, c("Wagners_Grade", "UTSA_Lavery_Grade")]

## vales in wagners grade
unique(cohen_data$Wagners_Grade)   # 1, 2, 3, 4, 5, NA

## vales in UTSA_Lavery_Grade
unique(cohen_data$UTSA_Lavery_Grade)  # I, II, III, NA
```

## Counts

```{r}
## Wagner Counts
count1_wag <- sum(cohen_data$Wagners_Grade == 1, na.rm = TRUE) #18
count2_wag <- sum(cohen_data$Wagners_Grade == 2, na.rm = TRUE) #26
count3_wag <- sum(cohen_data$Wagners_Grade == 3, na.rm = TRUE) #22
count4_wag <- sum(cohen_data$Wagners_Grade == 4, na.rm = TRUE) #14
count5_wag <- sum(cohen_data$Wagners_Grade == 5, na.rm = TRUE) #4

## Get levels the same
## Change I
cohen_data$UTSA_Lavery_Grade <- ifelse(cohen_data$UTSA_Lavery_Grade == "I", 1, cohen_data$UTSA_Lavery_Grade)
# Change II
cohen_data$UTSA_Lavery_Grade <- ifelse(cohen_data$UTSA_Lavery_Grade == "II", 2, cohen_data$UTSA_Lavery_Grade)
## Change III
cohen_data$UTSA_Lavery_Grade <- ifelse(cohen_data$UTSA_Lavery_Grade == "III", 3, cohen_data$UTSA_Lavery_Grade)

## UTSA Counts
countI_utsa <- sum(cohen_data$UTSA_Lavery_Grade == 1, na.rm = TRUE) #25
countII_utsa <- sum(cohen_data$UTSA_Lavery_Grade == 2, na.rm = TRUE) #18
countIII_utsa <- sum(cohen_data$UTSA_Lavery_Grade == 3, na.rm = TRUE) #41

## Probability that both match
#column totals

#probability 
p1wag=count1_wag/nrow(cohen_data)
p1utsa=countI_utsa/nrow(cohen_data)
p2wag=count2_wag/nrow(cohen_data)
p2utsa=countII_utsa/nrow(cohen_data)
p3wag=count3_wag/nrow(cohen_data)
p3utsa=countIII_utsa/nrow(cohen_data)

p1both=p1wag*p1utsa
p2both=p2wag*p2utsa
p3both=p3wag*p3utsa

p4=0
p5=0

## How many rows match
total_matching_rows <- sum(cohen_data$Wagners_Grade == cohen_data$UTSA_Lavery_Grade, 
                           na.rm = TRUE)  #47

## contingecny table values 
sum(cohen_data$Wagners_Grade == 1 & cohen_data$UTSA_Lavery_Grade == 1, na.rm=TRUE) #17



```

## Contengiency Table 

```{r Weighted Kappa}
## contingecny table values 
sum(cohen_data$Wagners_Grade == 5 & cohen_data$UTSA_Lavery_Grade == 3, na.rm=TRUE)

table = as.table(rbind(
  c(17, 1, 0, 0, 0), c(2, 12, 12, 0, 0), c(0, 4, 18, 0, 0), 
  c(4, 1, 9, 0, 0), c(2, 0, 2, 0, 0)))
categories = c("1", "2", "3",
                "4", "5")
dimnames(table) = list(Wagner = categories, UTSA = categories)
```

## Compute Kappa for two raters
```{r}

library(vcd)
(res.table = Kappa(table))

#Confidence interval
confint(res.table)


```

## Kappas equations

$P_0=\sum_{i=1}^{k} P_{ii}$
$P_e=\sum_{i=1}^{k} P_{i} P_{i}$
$k=\frac{P_0-P_e}{1-P_e}$

Here, weighted version is for ordinal data.

In our example, the Cohen’s kappa (k) = 0.651, which represents a fair to good strength of agreement according to Fleiss e al. (2003) classification. This is confirmed by the obtained p-value (p < 0.0001), indicating that our calculated kappa is significantly different from zero.


## Reserach Question
```{r Age NO CORR on wound}
####### AGE
age_wound <- df[, c("Age", "WoundScore_Total")]
    

# Using Spearman:
cor.test(age_wound$Age, age_wound$WoundScore_Total, alternative="greater",
         method="spearman", exact=T)

# Using Kendall:
cor.test(age_wound$WoundScore_Total, age_wound$Age, alternative="two.sided", 
         method="kendall", exact=T)

jonckheere.test(age_wound$WoundScore_Total,age_wound$Age,alternative="two.sided")
```

```{r in out NO Corr on Wound}
########## IN OUT PATIENT
inout_wound <- df[, c("In_Out_patient", "WoundScore_Total")]

jonckheere.test(age_wound$WoundScore_Total,age_wound$Age,alternative="two.sided")
```

```{r}

wound_goal<- df[, c("GoalScore_Total", "WoundScore_Total")]

# Using Pearson:
cor.test(wound_goal$GoalScore_Total, wound_goal$WoundScore_Total, alternative="greater", 
         method="pearson", exact=T)

# jonckheere.test
jonckheere.test(goal$GoalScore_Motivation, goal$GoalScore_Total, alternative="increasing")
# Using Spearman:
cor.test(goal$GoalScore_Motivation, goal$GoalScore_Total, alternative="greater",
         method="spearman", exact=T)
# Using kendall:
cor.test(goal$GoalScore_Motivation, goal$GoalScore_Total, alternative="greater", 
         method="kendall", exact=T)

```


```{r}
goal <- df[, c("GoalScore_Motivation", "GoalScore_Comprehension", "GoalScore_Compliance", "GoalScore_Support", "GoalScore_Insight", "GoalScore_Total")]
######### MOTIVATION
cor.test(goal$GoalScore_Motivation, goal$GoalScore_Total, alternative="greater", 
         method="pearson", exact=T)
######### CONPREHENSION
cor.test(goal$GoalScore_Comprehension, goal$GoalScore_Total, alternative="greater",
         method="pearson", exact=T)
######### Compliance
cor.test(goal$GoalScore_Compliance, goal$GoalScore_Total, alternative="greater", 
         method="pea", exact=T)
######### GoalScore_Support
cor.test(goal$GoalScore_Support, goal$GoalScore_Total, alternative="two.sided", 
         method="pearson", exact=T)
######### GoalScore_Insight
cor.test(goal$GoalScore_Insight, goal$GoalScore_Total, alternative="greater", 
         method="pearson", exact=T)
```

```{r}
## base model
df <- df[, c("GoalScore_Motivation", "GoalScore_Comprehension", "GoalScore_Compliance", "GoalScore_Insight", "WoundScore_Total")]
## non parametric regression
library(mgcv)
## Fit non-parametric regression model
model <- gam(WoundScore_Total ~ GoalScore_Motivation+GoalScore_Comprehension+GoalScore_Compliance+GoalScore_Insight, data=df)
## Print model summary
summary(model)
```

```{r}
# single predictor 
model2 <- gam(WoundScore_Total ~ GoalScore_Motivation, data=df)
summary(model2)
model3 <- gam(WoundScore_Total ~ GoalScore_Comprehension, data=df)
summary(model3)
model4 <- gam(WoundScore_Total ~ GoalScore_Compliance, data=df)
summary(model4)
model5 <- gam(WoundScore_Total ~ GoalScore_Insight, data=df)
summary(model5)
```

```{r}
# AIC
AIC(model2)
AIC(model3)
AIC(model4)
AIC(model5)
```








